### 2012-09-10 demo/taxize.Rmd
updated tutorial to recent changes...

# Install taxize (uncomment to run)
```{r}
# install_github("taxize_", "ropensci")
library(taxize); library(plyr); library(ape)
```

*********

# Phylomatic
## Fetch a phylogenetic tree from Phylomatic
```{r}
mytree <- "asteraceae/bidens/bidens_cernua
asteraceae/madia/madia_exigua
asteraceae/helianthus/helianthus_annuus"
```

## Using tree strings that need to be formatted within the function
### Return a tree in the phylo object format
```{r}
get_phylomatic_tree(mytree, 'TRUE', 'GET', 'xml', 'TRUE') 
```

### Return a tree in the phylo object format, using POST method
```{r}
get_phylomatic_tree(mytree, 'TRUE', 'POST', 'xml', 'TRUE')
```

### Return tree in newick format
```{r}
get_phylomatic_tree(mytree, 'TRUE', 'GET', 'new', 'FALSE')
```

### Get a tree in newick format, read using ape::read.tree, and plot
```{r phylog1 }
tree <- get_phylomatic_tree(mytree, 'TRUE', 'POST', 'new', 'FALSE')
treephylo <- read.tree(text = tree)
plot(treephylo)
```

*********

# Integrated Taxonomic Informaiton Service (ITIS)
## Search by term and search type
Note: TSN stands for Taxonomic Serial Number
```{r}
doc <- get_itis_xml("dolphin", "comname", "name")
doc <- get_itis_xml("grizzly bear", "comnameend", "name")
doc <- get_itis_xml("ursidae", "itistermssciname", "name")
```

## Take advantage of default arguments, spaces accepted, and display data.frame output
```{r}
doc <- get_itis_xml("Plethodon ")
head( parse_itis(doc) )
```

## Show higher taxonomy for a given TSN
```{r}
classification(685566, ID="tsn")
```

## Get a TSN from a sepcies name
```{r}
get_tsn(searchterm = "Quercus_douglasii", searchtype = "sciname")
```

*********

# Using ITIS and Phylomatic together 
## Use ITIS and get_phymat_format to format a string for each species to be submitted to Phylomatic
```{r phylog2}
dat_ <- laply(list("36616", "19322", "183327"), get_phymat_format, format='rsubmit')
dat_
dat_mine <- paste(dat_, collapse="%0D%0A") # collapse and replace \n's
dat_mine
tree <- get_phylomatic_tree(dat_mine, 'FALSE', 'GET', 'new', 'TRUE')
tree
plot(tree)
```

## A somewhat realistic workflow
### User's species list 
```{r}
splist <- c("annona cherimola", 'annona muricata', "quercus robur", 
						"shorea robusta", "pandanus patina", "oryza sativa", "durio zibethinus")
```

### Get TSN code for each species
```{r}
tsns <- get_tsn(splist, "sciname")
tsns
```

### Get Phylomatic formatted strings for each species
```{r}
dat_ <- laply(tsns, get_phymat_format, format = 'rsubmit')
dat_
```

### Send strings to Phylomatic and return newick tree
```{r}
tree <- get_phylomatic_tree(dat_, 'TRUE', 'GET', 'new', 'TRUE')
tree
```

### Plot tree, etc. 
```{r phylog3}
plot(tree)
```

*********

# Taxonomic Name Resolution Service
## Name matching using the TNRS
```{r}
tnrsmatch('best', taxnames = c('helianthus annuus', 'acacia', 'saltea'), output = 'names') # just best names
```

*********

# Global Names Resolver (GNR)
## Get just id's and names of sources in a data.frame
```{r highlight=TRUE}
tail( gnr_datasources(todf=T) )
```
***
## Give me the id for EOL (Encyclopedia of Life)
```{r highlight=TRUE}
out <- gnr_datasources(todf=T)
out[out$title == "EOL", "id"]
```
***
## Fuzzy search for sources with the word "zoo"
```{r highlight=TRUE}
out <- gnr_datasources(todf=T)
outdf <- out[agrep("zoo", out$title, ignore.case=T), ]
outdf[1:2,]
```
***
## Resolve some names
### Search for _Helianthus annuus_ and _Homo sapiens_, return a data.frame
```{r highlight=TRUE}
gnr(names = c("Helianthus annuus", "Homo sapiens"), returndf = TRUE)[1:2,]
```
***
### Search for the same species, with only using data source 12 (i.e., EOL)
```{r highlight=TRUE}
gnr(names = c("Helianthus annuus", "Homo sapiens"), data_source_ids="12", returndf = TRUE)
```

*********
# Tropicos: all Tropicos functions start with "tp_"
## 
```{r}
head(tp_getacceptednames(id = 25503923))
head(tp_getsynonyms(id = 25509881))
```

*********
# uBio
## Search uBio using the namebank_search API method.
```{r}
ubio_namebank_search(searchName = 'elephant', sci = 1, vern = 0)
head( ubio_namebank_search(searchName = 'Helianthus annuus', sci = 1, vern = 0)[,-c(2,3)] )
out <- lapply(list('Helianthus debilis','Astragalus aduncus'), function(x) ubio_namebank_search(searchName = x, sci = 1, vern = 0))
head( out[[2]][,-c(2,3)] ) # just Astragalus aduncus output
```