### 2013-02-12 inst/doc/taxize_ms.tex
added template for taxize manuscript in latex format

\documentclass[letterpaper,superscriptaddress,showkeys,longbibliography]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\hypersetup
{
    colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

\begin{document}

\title{taxize - taxonomic search and retrieval in R}

\author{Scott Chamberlain}
\email[E-mail: ]{myrmecocystus@gmail.com}
\affiliation{Biology Department, Simon Fraser University, Canada.}

\author{Eduard Sz\"{o}cs}
\email[E-mail: ]{szoe8822@uni-landau.de}
\affiliation{University Koblenz-Landau, Germany}

\keywords{taxonomy; R; sotware; data; API}

\maketitle

\section{Introduction}

asdfasfdf

\section{XXXXX}

asdfasfdf

\section{Conclusion}

asdfasfdf

\section{Funding}

SAC is supported by CANPOLIN of Canada. EZ is supported by XXXX.

\section{Acknowledgements}

asdfasfdf

\newpage
\bibliography{refs}

\end{document}