### 2013-04-16 inst/doc/taxize_tutorial.Rmd

# Install the development branch of taxize on github
```{r }
# install.packages("devtools")
# library(devtools)
# install_github("taxize_", "ropensci", "local_sql")
library(taxize)
library(plyr)
```

# Download the sqlite database from dropbox
https://www.dropbox.com/s/gz1vvsu2d0qps19/itis2_sqlite.zip

# Set path to database
Just set the path to your sqlite database once, then choose `locally = TRUE` in your ITIS function calls to use the local database. This creates an object in your options named "conn", which the ITIS functions call when `locally = TRUE`.
```{r}
taxize:::taxize_options(localpath = "~/github/ropensci/sql/itis2.sqlite") # change to your path
```

# Some examples
## Search by scientific name

### Single search 
#### Using the web API 
```{r}
searchbyscientificname("Tardigrada")
```

### Using local search
```{r}
searchbyscientificname("Tardigrada", locally=TRUE)
```

### As you can see, there is some performance improvement with even single queries
```{r}
system.time( searchbyscientificname("Tardigrada") )
system.time( searchbyscientificname("Tardigrada", locally=TRUE) )
```

### Search for multiple names

#### Using the web API, you have to submit one at a time using e.g., an lapply fxn
```{r}
spnames <- c("oryza sativa","Chironomus riparius","Helianthus annuus","Quercus lobata","Poa annua","Lampetra tridentata","Mordacia lapicida","Bathyraja abyssicola","Arhynchobatis asperrimus","Alytes obstetricans")
head( ldply(spnames, searchbyscientificname) )
```

#### Using local search, you can submit many names in one vector
```{r}
head( searchbyscientificname(srchkey=spnames, locally=TRUE) )
```

### Searches with many queries is where we see the major time savings 
```{r}
system.time( ldply(spnames, searchbyscientificname) ) # web API
system.time( searchbyscientificname(srchkey=spnames, locally=TRUE) ) # local sql search
```

## Get synonyms
```{r}
# many TSNs with the web API
ldply(c(202385,36616,19370,41107), getsynonymnamesfromtsn)

# Local search with sql
taxize:::taxize_options(localpath = "~/github/ropensci/sql/itis2.sqlite") # set to use your path
getsynonymnamesfromtsn(tsn = 526852, locally=TRUE) # a single TSN
getsynonymnamesfromtsn(tsn=c(202385,36616,19370,41107), locally=TRUE) # many TSNs

# any you can return the index on the data.frame, that is, the TSN you searched with for easy parsing afterwards
getsynonymnamesfromtsn(tsn=c(202385,36616,19370,41107), locally=TRUE, returnindex=TRUE)
```

# Here's what the one of the functions looks like with the sql syntax
```{r eval=FALSE }
gettsnbyvernacularlanguage <- function(language = NA, locally = FALSE) 
{
  if (locally) {
    sqlconn <- getOption("conn")
    query_TSNS_BY_LANGUAGE <- paste("select", 
        paste("CASE", paste(sapply(language, function(x) 
        paste("WHEN language LIKE ", paste("'", x, "'", sep = ""), " THEN ", 
        paste0("'",x,"'"), sep = ""), USE.NAMES=FALSE),collapse=" "),"END AS querystring,"),
        "tsn, vernacular_name from vernaculars where ", paste0(" language like ", 
        sapply(language, function(x) paste("'", x, "'", sep = ""), USE.NAMES=FALSE), collapse=" OR "), 
        " order by tsn, vernacular_name;")
    temp <- dbGetQuery(conn=sqlconn, query_TSNS_BY_LANGUAGE)
    return(data.frame(language = temp$querystring, comname = temp$vernacular_name, tsn = temp$tsn))
  }
  else {
    url = "http://www.itis.gov/ITISWebService/services/ITISService/getTsnByVernacularLanguage"
    args <- list()
    if (!is.na(language)) 
      args$language <- language
    tt <- getForm(url, .params = args, ..., curl = curl)
    out <- xmlParse(tt)
    namespaces <- c(namespaces <- c(ax21 = "http://data.itis_service.itis.usgs.gov/xsd"))
    nodes <- getNodeSet(out, "//ax21:commonName", namespaces = namespaces)
    comname <- sapply(nodes, xmlValue)
    nodes <- getNodeSet(out, "//ax21:tsn", namespaces = namespaces)
    tsn <- sapply(nodes, xmlValue)
    data.frame(comname = comname, tsn = tsn)
  }
}
```

#### Note:
Not all function that use ITIS data have a local sql option yet, but many do.  Send any bug reports [here](https://github.com/ropensci/taxize_/issues)