### 2013-04-27 inst/doc/taxize_withcode.Rnw
updated the mansucript

\documentclass[letterpaper,superscriptaddress,showkeys,longbibliography]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
\bibliographystyle{ecol_let}
\hypersetup
{
  colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
## set global chunk options
opts_chunk$set(fig.path='figure/', fig.align='center', fig.show='hold', par=TRUE, warning=FALSE, message=FALSE, error=FALSE, comment=NA)
@

\title{taxize - taxonomic search and retrieval in R}

\author{Scott Chamberlain}
\email[E-mail: ]{myrmecocystus@gmail.com}
\affiliation{Biology Department, Simon Fraser University, Canada.}

\author{Eduard Sz\"{o}cs}
\email[E-mail: ]{szoe8822@uni-landau.de}
\affiliation{University Koblenz-Landau, Germany}
\keywords{taxonomy; R; sotware; data; API}

\maketitle

\section{Introduction}

some text here

\section{The case for taxize}

There are a large suite of applications developed around the problem of searching for, resolving, and getting higher taxonomy for species names. For example, Linnaeus \url{http://linnaeus.sourceforge.net/} provides ability to search for taxonomic names in documents and normalize names. In addition, there are many web interfaces to search for and normalize names such as Encyclopedia of Life's Global Names Resolver \url{http://resolver.globalnames.org/}, uBio tools \url{http://www.ubio.org/index.php?pagename=sample_tools}, and iPlant's Taxonomic Name Resolution Service \url{http://tnrs.iplantcollaborative.org/}.

All of these tools provide great ways to search for taxonomic names and resolve them in some cases. However, scientists ideally need a tool that can be used programmatically, and thus be made reproducible, and highly customizeable. The goal of taxize is to make it easy to create reproducible and easy to use workflows for searching for taxonomic names, resolving them, getting higher taxonomic names, and other tasks related to research dealing with species.

\section{Data sources}

taxize uses many data sources, and more can easily be added.

\begin{table}[ht]
\caption{Data sources used in taxize} % title of Table
\centering % used for centering table
\begin{tabular}{|l|ccc|c|}
\hline
Source name &  Name search & Name resolution & Phylogeny & URL  \\
\hline
Encyclopedia of Life & Yes & See GNR below & No & \url{http://eol.org/} \\
Integrated Taxonomic Resolution Service & Yes & Synonyms & No & X\\
iPlant Taxonomic Name Resolution Service & Yes & Yes & No & X \\
Phylomatic & No & X & No & X \\
uBio & Yes & X & No & X \\
Global Names Resolver & Yes & X & No & X \\
Global Names Index & Yes & X & No & X \\
IUCN Red List & Yes & X & No & X \\
Tropicos & Yes & X & No & X \\
Plantminer & Yes & X & No & X \\
Theplantlist.org & Yes & X & No & X \\
Catalogue of Life & Yes & X & No & X \\
National Center for Biotechnology Information (NCBI) & Yes & X & No & X \\
\hline
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

\section{Use cases}

There are a variety of use cases for which taxize is ideally suited, and few side cases in which taxize can be useful. We discuss five ideal use cases for taxize at length, and highlight the side cases in brief.

\subsection{Installing taxize}

First, let's install taxize. There are two versions of taxize, a stable release that can be installed from the R package repository, CRAN, and from GitHub, where the code is developed.

Installing from CRAN or GitHub

<<installtaxize, eval=FALSE>>=
## From CRAN
install.packages("taxize")

## From GitHub
install_github("taxize_", "ropensci")
@

Loading into your R session

<<loadtaxize, eval=TRUE>>=
library(taxize)
@

\subsection{Resolve taxonomic names}

This is a common task in biology. We often have a list of species names and we want to know if a) we have the most up to date names, b) our names are spelled correctly, and c) if we have common names, we likely need the scientific names. One way to resolve names is via the Global Names Resolver service provided by the Encyclopedia of Life (\url{http://resolver.globalnames.org/}).

<<resolvenames, eval=TRUE>>=
# Here, we are searching for two misspelled names
temp <- gnr_resolve(names = c("Helianthos annus", "Homo saapiens"), returndf = TRUE)

# let's take a peek at the data, excluding the data source ID and score columns
temp[,-c(1,4)]
@

Looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}, cool!

Another approach is using the Taxonomic Name Resolution Service via the Taxosaurus API (\url{http://taxosaurus.org/}).

<<tnrs, eval=TRUE>>=
# Lets set our list of species names
mynames <- c("Helianthus annuus", "Pinus contort", "Poa anua", "Abis magnifica",
		"Rosa california", "Festuca arundinace", "Sorbus occidentalos","Madia sateva")

# And we'll call the API with the tnrs function
tnrs(query = mynames)[,-c(5:7)]
@

It looks like there are a few corrections: e.g., \emph{Madia sateva} should be \emph{Madia sativa}, and \emph{Rosa california} should be \emph{Rosa californica}.

\subsection{Retrieve higher taxonomic names}

Another task biologists often face is wanting to get higher taxonomic names for their list of taxa. If you have the higher taxonomy you can put in to context the relationships of your list (i.e., Species A and B are in Family X), as opposed to not knowing that Species A and B are closely related. A number of data sources provide this type of capability. First, let's take a look at the Integrated Taxonomic Information Service (ITIS).

<<itisclass, eval=TRUE>>=
specieslist <- c("Abies procera","Pinus contorta")
classification(get_tsn(specieslist, "sciname"))
@

It turns out both species are in the family Pinaceae. You can also get this type of information from the NCBI by doing \emph{classification(get\_uid(specieslist))}

Instead of a full classification, you may only wan a single name, say a family name for your species of interest. The function \emph{tax_name} is built just for this purpose. And you can specify the data source you retrieve the taxonomic name from with the \emph{db} parameter.

<<taxname, eval=TRUE>>=
tax_name(query="Helianthus annuus", get="family", db="itis")
tax_name(query="Helianthus annuus", get="family", db="ncbi")
@

\subsection{Retrieve a phylogeny}

There is an increasingly common use case: many biologists are not adequately trained in reconstructing phylogenies. However, so-called \emph{taxonomic phylogenies} can be constructed from simply knowing the higher taxonomic classification of a set of taxa. There are few taxon groups for which we can get phylogenies simply based on taxonomy; one of these is for angiosperms, called Phylomatic \cite{webb2005}. We have created a workflow in taxize that lets you input a simple species list, and then taxize does work behind the scenes to get higher taxonomic names, which are required by Phylomatic to get a phylogeny. Here is a short example.

<<phylomaticphylogeny, eval=TRUE, dev='png', out.width='.8\\linewidth', out.height='3in'>>=
library(doMC)
registerDoMC(cores=4)

# input the taxonomic names
taxa <- c("Poa annua", "Abies procera", "Helianthus annuus")

# fetch the tree - the formatting of names and higher taxonmy is done within the function
tree <- phylomatic_tree(taxa=taxa, get = 'POST', informat='newick', method = "phylomatic", storedtree = "R20120829", taxaformat = "slashpath", outformat = "newick", clean = "true")

tree$tip.label <- capwords(tree$tip.label)

# plot the tree
plot(tree, cex=1.2)
@

\subsection{What taxa are in children of my taxon of interest?}

If you aren't a taxonomic specialist on a particular taxon you likely don't know what children taxa are within a family, or within a genus. You can of course go to a website like Wikispecies (\url{http://species.wikimedia.org/wiki/Main_Page}) or Encyclopedia of Life (\url{http://eol.org/}). taxize provides an easy way for you to search for downstream taxa, both for the Catalogue of Life (CoL; \url{http://www.catalogueoflife.org/}) and the Integrated Taxanomic Information Database (\url{http://www.itis.gov/}). Here is a short example using the CoL in which we want to find all the species within the genus \emph{Apis} (honey bees).

<<downstream, eval=TRUE>>=
col_downstream(name="Apis", downto="Species")[[1]]
@

\subsection{IUCN Status}

There are a number of things we can do once we have the correct taxonomic names. One thing we can do is ask about the conservation status of a species. We have provided a set of functions, \emph{iucn\_summary} and \emph{iucn\_status}, to search for species names, and extract the status information, respectively. Here, we search for the Panther and Lynx.

<<XXXX, eval=TRUE>>=
ia <- iucn_summary(c("Panthera uncia", "Lynx lynx"))
iucn_status(ia)
@

It turns out that the Panther is endangered (EN) and the Lynx is of least concern (LN).

\section{Conclusion}

some text here

\section{Funding}

SAC is supported by CANPOLIN of Canada. EZ is supported by XXXX.

\section{Acknowledgements}

some text here

\newpage
\bibliography{refs}

\end{document}