### 2013-05-29 inst/doc/taxize_withcode_appB.Rnw

\section{Matching species tables with different taxonomic resolution}

Trait-based approaches are a promising tool in ecology. Unlike taxonomy-based methods, traits may not be constrained to biogeographic boundaries \citep{baird_toward_2011} and have potential to disentangle the effects of multiple stressors \citep{statzner_can_2010}.

To analyse trait-composition abundance data must be matched with trait databases like \citet{usseglio-polatera_biological_2000}. However these two datatables may contain species information on different taxonomic levels and perhabs data must be aggregated to a joint taxomic level.

taxize can help in this data-cleaning step, providing a reproducible workflow. Here we illustrate this on a small fictitious example.

Suppose we have fuzzy coded trait table with 2 traits with 3 respectively 2 modalities:
<<>>=
(traits <- read.table(header = TRUE, sep = ';', text = 'taxon;T1M1;T1M2;T1M3;T2M1;T2M2
Gammarus sp.;0;0;3;1;3
Potamopyrgus antipodarum;1;0;3;1;3
Coenagrion sp.;3;0;1;3;1
Enallagma cyathigerum;0;3;1;0;3
Erythromma sp.;0;0;3;3;1
'))
@

And want to match this to a table with abundances:
<<>>=
(abundances <- read.table(header = TRUE, sep = ';', text = 'taxon;abundance
Gammarus roeseli;5
Gammarus roeseli;6
Gammarus tigrinus;7
Gammarus tigrinus;8
Coenagrionidae;10
Coenagrionidae;6
Potamopyrgus antipodarum;10
'))
@