### 2013-06-04 inst/doc/taxize_withcode_appB.Rnw
More work on AppB

\section{Matching species tables with different taxonomic resolution}

Trait-based approaches are a promising tool in ecology. Unlike taxonomy-based methods, traits may not be constrained to biogeographic boundaries \citep{baird_toward_2011} and have potential to disentangle the effects of multiple stressors \citep{statzner_can_2010}.

To analyse trait-composition abundance data must be matched with trait databases like \citet{usseglio-polatera_biological_2000}. However these two datatables may contain species information on different taxonomic levels and perhabs data must be aggregated to a joint taxomic level.

taxize can help in this data-cleaning step, providing a reproducible workflow. Here we illustrate this on a small fictitious example.

Suppose we have fuzzy coded trait table with 2 traits with 3 respectively 2 modalities:
<<traitdata>>=
(traits <- read.table(header = TRUE, sep = ';', stringsAsFactors=FALSE, text = 'taxon;T1M1;T1M2;T1M3;T2M1;T2M2
Gammarus sp.;0;0;3;1;3
Potamopyrgus antipodarum;1;0;3;1;3
Coenagrion sp.;3;0;1;3;1
Enallagma cyathigerum;0;3;1;0;3
Erythromma sp.;0;0;3;3;1
Baetis sp.;0;0;0;0;0
'))
@

And want to match this to a table with abundances:
<<abundancedata>>=
(abundances <- read.table(header = TRUE, sep = ';', stringsAsFactors=FALSE, text = 'taxon;abundance;sample
Gammarus roeseli;5;1
Gammarus roeseli;6;2
Gammarus tigrinus;7;1
Gammarus tigrinus;8;2
Coenagrionidae;10;1
Coenagrionidae;6;2
Potamopyrgus antipodarum;10;1
xxxxx;10;2
'))
@

First we do some basic data-cleaning and create a looktable, that will link taxa in trait table with taxa in the abundance table.
<<cleaning>>=
# first we remove ' sp.' from out trait table:
traits$taxon_cleaned <- tolower(gsub(' sp.', '', traits$taxon))

# since abundance tables can be very long with repeating taxa, we look only at unique
# taxon names 
# This will be a lookup-table linking taxon names between both tables
lookup <- data.frame(taxon = tolower(unique(abundances$taxon)), stringsAsFactors=FALSE)
@

The we query the taxonomic hierarchy for both tables, since this is the backbone of this procedure:
<<get_classification, cache=TRUE>>=
require(taxize)
traits_classi <- classification(get_uid(traits$taxon_cleaned))
lookup_classi <- classification(get_uid(lookup$taxon))
@

First we look if we get any direct matches between taxon names:
<<direct_matches>>=
# first search for direct matches
direct <- match(lookup$taxon, traits$taxon_cleaned)
# and add the matched name to our lookup table
lookup$traits <- tolower(traits$taxon[direct])
lookup$type <- ifelse(!is.na(direct), "direct", NA)
lookup
@

We found a direct match - \emph{potamopyrgus antipodarum} - so nothing to do here.

Next we look for species which are on a better taxonomic resolution than our in our trait table. 
For these species we'll use the trait-data since no better is available.

<<lower>>=
# look for cases where taxonomic resolution in abundance data is higher than in trait data:
# here we take the trait-values for the lower resolution
for(i in which(is.na(lookup$traits))){
  if(is.data.frame(lookup_classi[[i]])){
    matches <- tolower(lookup_classi[[i]]$ScientificName) %in% traits$taxon_cleaned
    if(any(matches)){
      lookup$traits[i] <- tolower(lookup_classi[[i]]$ScientificName[matches])
      lookup$type[i] <- lookup_classi[[i]]$Rank[matches]
    }
  }
}
lookup
@

So our abundance data has two Gammarus species, however trait data is only on Genus level.

Our next step is to search for species were we have to aggregate tait-data, since our abundance data is on a lower taxonomic level.
We are walking the taxomomic latter for the species in our trait-data upwards and search for matches with out abundance data. Since we'll have many taxa in the trait-data belonging to one taxon, we'll take the median modality scores as an approximation. Of course also other methods may be used here, e.g. weighting by genetic distance.

<<>>=
# look for cases taxonomic resolution in abundance data is lower than in trait data,
# here we need to aggregate the trait-values (eg. median value for modality)

for(i in which(is.na(lookup$traits))){
  # find matches
  agg <- sapply(traits_classi, function(x) any(tolower(x$ScientificName) %in% lookup$taxon[i]))
  if(sum(agg) > 1){
    # add taxon as aggregate to trait-table
    traits <- rbind(traits, c(paste0(lookup$taxon[i], '-aggregated'), apply(traits[agg, 2:6], 2, median), paste0(lookup$taxon[i], '-aggregated')))
    # fill lookup table
    lookup$traits[i] <- paste0(lookup$taxon[i], '-aggregated')
    lookup$type[i] <- 'aggregated'
  }
}
lookup
@

Finally we have one taxon left. We remove this from our dataset:
<<>>=
abundances <- abundances[!abundances$taxon == lookup$taxon[is.na(lookup$traits)], ]
@

No we can create species x sites and traits x species matrices, which could be plugged into methods to analyse trait responses \citep{kleyer_assessing_2012}.

<<>>=
# species (as matched with trait table) by site matrix
abundances$traits_taxa <- lookup$traits[match(tolower(abundances$taxon), lookup$taxon)]

require(reshape2)
# reshape data to long format and name rows by samples
L <- dcast(abundances, sample ~ traits_taxa, fun.aggregate=sum, value.var = 'abundance')
rownames(L) <- L$sample
L$sample <- NULL
L

# traits by species matrix
Q <- traits[, 2:7][match(names(L), traits$taxon_cleaned), ]
rownames(Q) <- Q$taxon_cleaned
Q$taxon_cleaned <- NULL
Q

# check
all(rownames(Q) == colnames(L))
@