### 2013-08-26 inst/doc/taxize_infotable.Rmd
added draft vignettes

<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->

Some key functions in taxize, what they do, and their data sources.
======

Function name | What it does | Source |
----------- | ----------- | ----------- |
apg_lookup | Changes names to match the APGIII list | Angiosperm Phylogeny Group |
classification | Upstream classification | Various |
col_children | Direct children | Catalogue of Life |
col_downstream | Downstream taxa to specified rank | Catalogue of Life |
eol_hierarchy | Upstream classification | Encyclopedia of Life |
eol_search | Search EOL taxon information | Encyclopedia of Life |
get_seqs | Get NCBI sequences | National Center for Biotechnology Information |
get_tsn | Get ITIS TSN | Integrated Taxonomic Information System |
get_uid | Get NCBI UID | National Center for Biotechnology Information |
searchbycommonname | Search ITIS by common name | Integrated Taxonomic Information System |
searchbyscientificname | Search ITIS by scientific name | Integrated Taxonomic Information System |
gisd_isinvasive | Invasiveness status | Global Invasive Species Database |
gni_parse | Parse scientific names into components | Global Names Index |
gni_search | Search EOL's global names index | Global Names Index |
gnr_resolve | Resolve names using EOL's global names index | Global Names Resolver |
itis_downstream | Downstream taxa to specified rank | Integrated Taxonomic Information System |
iucn_status | IUCN status | IUCN Red List |
phylomatic_tree | Get a plant Phylogeny | Phylomatic |
plantminer | Search Plantminer | Plantminer |
tax_name | Get taxonomic name for specific rank | Various |
tax_rank | Get rank of a taxonomic name | Various |
tnrs | Resolve names using iPlant | iPlant Taxonomic Name Resolution Service |
tp_acceptednames | Check for accepted names using Tropicos | Tropicos |
tpl_search | Search the Plant List | The Plant List |
ubio_namebank | Search uBio | uBio |