### 2014-01-31 inst/vign/taxize_infotable.Rmd
changed vignettes to not build during pkg install, new makefile

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Key functions}
-->

Some key functions in taxize, what they do, and their data sources.
======

Function name | What it does | Source |
----------- | ----------- | ----------- |
classification | Upstream classification | Various |
col_children | Direct children | Catalogue of Life |
col_downstream | Downstream taxa to specified rank | Catalogue of Life |
eol_hierarchy | Upstream classification | Encyclopedia of Life |
eol_search | Search EOL taxon information | Encyclopedia of Life |
ncbi_search | Get NCBI sequences | National Center for Biotechnology Information |
get_ids | Get taxonomic identifiers | Five sources |
comm2sci | Get scientific from common names | EOL, NCBI, ITIS, Tropicos |
sci2comm | Get common from scientific names | EOL, NCBI, ITIS |
gisd_isinvasive | Invasiveness status | Global Invasive Species Database |
gni_parse | Parse scientific names into components | Global Names Index |
gni_search | Search EOL's global names index | Global Names Index |
gnr_resolve | Resolve names using EOL's global names index | Global Names Resolver |
itis_downstream | Downstream taxa to specified rank | Integrated Taxonomic Information System |
iucn_summary | Get IUCN status for a species | IUCN Red List |
phylomatic_tree | Get a plant Phylogeny | Phylomatic |
plantminer | Search Plantminer | Plantminer |
tax_name | Get taxonomic name for specific rank | Various |
tax_rank | Get rank of a taxonomic name | Various |
tnrs | Resolve names using Taxosaurus/iPlant | Taxosaurus Taxonomic Name Resolution Service |
tp_accnames | Check for accepted names using Tropicos | Tropicos |
ubio_namebank | Search uBio | uBio |