### 2016-07-22 inst/vign/name_cleaning.Rmd
#549 name cleaning vignette work

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Strategies for programmatic name cleaning}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

Strategies for programmatic name cleaning
=========================================

`taxize` offers interactive prompts when using `get_*()` functions (e.g., `get_tsn()`).
These prompts make it easy in interactive use to select choices when there are more
than one match found.

However, to make your code reproducible you don't want interactive prompts.

This vignette covers various options for programmatic name cleaning.

```{r}
library("taxize")
```

## get_* functions

When using `get_*()` functions programatically, you have a few options.

### rows parameter

Normally, if you get more than one result, you get a prompt asking you
to select which taxon you want.

```{r eval=FALSE}
get_tsn(searchterm = "Quercus b")
#>       tsn                       target             commonnames    nameusage
#> 1   19298             Quercus beebiana                         not accepted
#> 2  507263       Quercus berberidifolia               scrub oak     accepted
#> 3   19300              Quercus bicolor         swamp white oak     accepted
#> 4   19303             Quercus borealis                         not accepted
#> 5  195131 Quercus borealis var. maxima                         not accepted
#> 6  195166            Quercus boyntonii Boynton's sand post oak     accepted
#> 7  506533              Quercus brantii             Brant's oak     accepted
#> 8  195150            Quercus breviloba                         not accepted
#> 9  195099              Quercus breweri                         not accepted
#> 10 195168             Quercus buckleyi               Texas oak     accepted
#>
#> More than one TSN found for taxon 'Quercus b'!
#>
#>             Enter rownumber of taxon (other inputs will return 'NA'):
#>
#> 1:
```

Instead, we can use the rows parameter to specify which records we want
by number only (not by a name itself). Here, we want the first 3 records:

```{r eval=FALSE}
get_tsn(searchterm = 'Quercus b', rows = 1:3)
#>     tsn           target     commonnames    nameusage
#> 1 19298 Quercus beebiana                 not accepted
#> 2 19300  Quercus bicolor swamp white oak     accepted
#> 3 19303 Quercus borealis                 not accepted
#>
#> More than one TSN found for taxon 'Quercus b'!
#>
#>             Enter rownumber of taxon (other inputs will return 'NA'):
#>
#> 1:
```

However, you still get a prompt as there is more than one result.

Thus, for full programmatic usage, you can specify a single row, if you happen
to know which one you want:

```{r}
get_tsn(searchterm = 'Quercus b', rows = 3)
```

In reality it is unlikely you'll know which row you want, unless perhaps you 
just one one result from each query, regardless of what it is.

### underscore methods

A better fit for programmatic use are underscore methods. Each `get_*()` function
has a sister method with and trailing underscore, e.g., `get_tsn()` and `get_tsn_()`.

```{r}
get_tsn_(searchterm = "Quercus b")
```

The result is a single data.frame for each taxon queried, which can be 
processed downstream with whatever logic is required in your workflow.

You can also combine `rows` parameter with underscore functions, as a single
number of a range of numbers:

```{r}
get_tsn_(searchterm = "Quercus b", rows = 1)
```

```{r}
get_tsn_(searchterm = "Quercus b", rows = 1:2)
```

## gnr_resolve

Some functions in `taxize` are meant specifically for name cleaning. One of those
is `gnr_resolve()`.

`gnr_resolve()` doesn't provide prompts as do `get_*()` functions, but instead
return data.frame's. So we don't face the same problem, and can use `gnr_resolve()` 
in a programmatic workflow straight away.

```{r}
spp <- names_list(rank = "species", size = 10)
gnr_resolve(names = spp, preferred_data_sources = 11)
```