### 2020-09-17 README.Rmd
update readme, move contribs list in redame to txt file to avoid cran url check errors

taxize
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![cran checks](https://cranchecks.info/badges/worst/taxize)](https://cranchecks.info/pkgs/taxize)
[![R-CMD-check](https://github.com/ropensci/taxize/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/taxize/actions/)
[![codecov](https://codecov.io/gh/ropensci/taxize/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/taxize)
[![rstudio mirror downloads](https://cranlogs.r-pkg.org/badges/taxize)](https://github.com/r-hub/cranlogs.app)
[![cran version](https://www.r-pkg.org/badges/version/taxize)](https://cran.r-project.org/package=taxize)

`taxize` allows users to search over many taxonomic data sources for species names (scientific and common) and download up and downstream taxonomic hierarchical information - among other things.

The taxize book: https://taxize.dev

Package documentation: https://docs.ropensci.org/taxize/

## Data sources currently implemented

<table>
<colgroup>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
<col style="text-align:left;"/>
</colgroup>

<thead>
<tr>
  <th style="text-align:left;">Souce</th>
	<th style="text-align:left;">Function prefix</th>
	<th style="text-align:left;">API Docs</th>
	<th style="text-align:left;">API key</th>
</tr>
</thead>

<tbody>
<tr>
	<td style="text-align:left;">Encylopedia of Life</td>
	<td style="text-align:left;"><code>eol</code></td>
	<td style="text-align:left;"><a href="https://eol.org/docs/what-is-eol/data-services">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Taxonomic Name Resolution Service</td>
	<td style="text-align:left;"><code>tnrs</code></td>
	<td style="text-align:left;">none</td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Integrated Taxonomic Information Service</td>
	<td style="text-align:left;"><code>itis</code></td>
	<td style="text-align:left;"><a href="https://www.itis.gov/ws_description.html">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Global Names Resolver</td>
	<td style="text-align:left;"><code>gnr</code></td>
	<td style="text-align:left;"><a href="http://resolver.globalnames.org/api">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Global Names Index</td>
	<td style="text-align:left;"><code>gni</code></td>
	<td style="text-align:left;"><a href="https://github.com/dimus/gni/wiki/api">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">IUCN Red List</td>
	<td style="text-align:left;"><code>iucn</code></td>
	<td style="text-align:left;"><a href="https://apiv3.iucnredlist.org/api/v3/docs">link</a></td>
	<td style="text-align:left;"><a href="https://apiv3.iucnredlist.org/api/v3/token">link</a></td>
</tr>
<tr>
	<td style="text-align:left;">Tropicos</td>
	<td style="text-align:left;"><code>tp</code></td>
	<td style="text-align:left;"><a href="http://services.tropicos.org/help">link</a></td>
	<td style="text-align:left;"><a href="http://services.tropicos.org/help?requestkey">link</a></td>
</tr>
<tr>
	<td style="text-align:left;">Theplantlist dot org</td>
	<td style="text-align:left;"><code>tpl</code></td>
	<td style="text-align:left;">**</td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">National Center for Biotechnology Information</td>
	<td style="text-align:left;"><code>ncbi</code></td>
	<td style="text-align:left;">none</td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">CANADENSYS Vascan name search API</td>
	<td style="text-align:left;"><code>vascan</code></td>
	<td style="text-align:left;"><a href="https://data.canadensys.net/vascan/api">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">International Plant Names Index (IPNI)</td>
	<td style="text-align:left;"><code>ipni</code></td>
	<td style="text-align:left;">none</td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Barcode of Life Data Systems (BOLD)</td>
	<td style="text-align:left;"><code>bold</code></td>
	<td style="text-align:left;"><a href="http://www.boldsystems.org/index.php/Resources">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">National Biodiversity Network (UK)</td>
	<td style="text-align:left;"><code>nbn</code></td>
	<td style="text-align:left;"><a href="https://data.nbn.org.uk/Documentation/Web_Services/Web_Services-REST/resources/restapi/rest.html">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Index Fungorum</td>
	<td style="text-align:left;"><code>fg</code></td>
	<td style="text-align:left;">none</td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">EU BON</td>
	<td style="text-align:left;"><code>eubon</code></td>
	<td style="text-align:left;"><a href="https://cybertaxonomy.eu/eubon-utis/doc.html">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Index of Names (ION)</td>
	<td style="text-align:left;"><code>ion</code></td>
	<td style="text-align:left;"><a href="http://www.organismnames.com/">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Open Tree of Life (TOL)</td>
	<td style="text-align:left;"><code>tol</code></td>
	<td style="text-align:left;"><a href="https://github.com/OpenTreeOfLife/germinator/wiki/Open-Tree-of-Life-Web-APIs">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">World Register of Marine Species (WoRMS)</td>
	<td style="text-align:left;"><code>worms</code></td>
	<td style="text-align:left;"><a href="https://www.marinespecies.org/aphia.php?p=webservice">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">NatureServe</td>
	<td style="text-align:left;"><code>natserv</code></td>
	<td style="text-align:left;"><a href="https://services.natureserve.org/BrowseServices/getSpeciesData/getSpeciesListREST.jsp">link</a></td>
	<td style="text-align:left;"><a href="https://services.natureserve.org/developer/index.jsp">link</a></td>
</tr>
<tr>
	<td style="text-align:left;">Wikipedia</td>
	<td style="text-align:left;"><code>wiki</code></td>
	<td style="text-align:left;"><a href="https://www.mediawiki.org/wiki/API:Main_page">link</a></td>
	<td style="text-align:left;">none</td>
</tr>
<tr>
	<td style="text-align:left;">Kew's Plants of the World</td>
	<td style="text-align:left;"><code>pow</code></td>
	<td style="text-align:left;">none</td>
	<td style="text-align:left;">none</td>
</tr>
</tbody>
</table>

**: There are none! We suggest using `TPL` and `TPLck` functions in the [taxonstand package](https://cran.r-project.org/package=Taxonstand). We provide two functions to get bulk data: `tpl_families` and `tpl_get`.

\***: There are none! The function scrapes the web directly.

### May be in taxize in the future...

See the [datasources](https://github.com/ropensci/taxize/labels/datasources) tag in the issue tracker

<br>

## Installation

### Stable version from CRAN

```{r eval=FALSE}
install.packages("taxize")
```

### Development version from GitHub

Windows users install Rtools first.

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("ropensci/taxize")
```

```{r}
library('taxize')
```

```{r echo=FALSE}
taxize_options(FALSE, quiet = TRUE)
```

## Get unique taxonomic identifier from NCBI

Alot of `taxize` revolves around taxonomic identifiers. Because, as you know, names can be a mess (misspelled, synonyms, etc.), it's better to get an identifier that a particular data source knows about, then we can move forth acquiring more fun taxonomic data.

```{r get_uid}
uids <- get_uid(c("Chironomus riparius", "Chaetopteryx"))
```

## Retrieve classifications

Classifications - think of a species, then all the taxonomic ranks up from that species, like genus, family, order, class, kingdom.

```{r classification}
out <- classification(uids)
lapply(out, head)
```

## Immediate children

Get immediate children of _Salmo_. In this case, _Salmo_ is a genus, so this gives species within the genus.

```{r children}
children("Salmo", db = 'ncbi')
```

## Downstream children to a rank

Get all species in the genus _Apis_

```{r downstream}
downstream(as.tsn(154395), db = 'itis', downto = 'species', mesages = FALSE)
```

## Upstream taxa

Get all genera up from the species _Pinus contorta_ (this includes the genus of the species, and its co-genera within the same family).

```{r upstream}
upstream("Pinus contorta", db = 'itis', upto = 'Genus', mesages = FALSE)
```

## Get synonyms

```{r synonyms}
synonyms("Acer drummondii", db="itis")
```

## Get taxonomic IDs from many sources

```{r get_ids}
get_ids("Salvelinus fontinalis", db = c('itis', 'ncbi'), mesages = FALSE)
```

You can limit to certain rows when getting ids in any `get_*()` functions

```{r get_ids2}
get_ids("Poa annua", db = "gbif", rows=1)
```

Furthermore, you can just back all ids if that's your jam with the `get_*_()` functions (all `get_*()` functions with additional `_` underscore at end of function name)

```{r get_ids_}
get_ids_(c("Chironomus riparius", "Pinus contorta"), db = 'nbn', rows=1:3)
```

## Common names from scientific names

```{r sci2comm}
sci2comm('Helianthus annuus', db = 'itis')
```

## Scientific names from common names

```{r comm2sci}
comm2sci("black bear", db = "itis")
```

## Lowest common rank among taxa

```{r lowest}
spp <- c("Sus scrofa", "Homo sapiens", "Nycticebus coucang")
lowest_common(spp, db = "ncbi")
```

## Coerce codes to taxonomic id classes

`numeric` to `uid`

```{r as.uid}
as.uid(315567)
```

`list` to `uid`

```{r as.uid2}
as.uid(list("315567", "3339", "9696"))
```

## Coerce taxonomic id classes to a data.frame

```{r as.uid3}
out <- as.uid(c(315567, 3339, 9696))
(res <- data.frame(out))
```

## Screencast

<a href="https://vimeo.com/92883063"><img src="man/figures/screencast.png" width="400"></a>

## Contributing

See our [CONTRIBUTING](https://github.com/ropensci/taxize/blob/master/.github/CONTRIBUTING.md) document.

## Road map

Check out our [milestones](https://github.com/ropensci/taxize/milestones) to see what we plan to get done for each version.

## Meta

* Please [report any issues or bugs](https://github.com/ropensci/taxize/issues).
* License: MIT
* Get citation information for `taxize` in R doing `citation(package = 'taxize')`
* Please note that this package is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

[![rofooter](https://ropensci.org/public_images/github_footer.png)](https://ropensci.org)