### 2020-09-18 README.Rmd
bump version after cran push, add readme contribs back to readme

taxize
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![cran checks](https://cranchecks.info/badges/worst/taxize)](https://cranchecks.info/pkgs/taxize)
[![R-CMD-check](https://github.com/ropensci/taxize/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/taxize/actions/)
[![codecov](https://codecov.io/gh/ropensci/taxize/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/taxize)
[![rstudio mirror downloads](https://cranlogs.r-pkg.org/badges/taxize)](https://github.com/r-hub/cranlogs.app)
[![cran version](https://www.r-pkg.org/badges/version/taxize)](https://cran.r-project.org/package=taxize)

`taxize` allows users to search over many taxonomic data sources for species names (scientific and common) and download up and downstream taxonomic hierarchical information - among other things.

The taxize book: https://taxize.dev

Package documentation: https://docs.ropensci.org/taxize/

## Installation

### Stable version from CRAN

```{r eval=FALSE}
install.packages("taxize")
```

### Development version from GitHub

Windows users install Rtools first.

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("ropensci/taxize")
```

```{r}
library('taxize')
```

## Screencast

<a href="https://vimeo.com/92883063"><img src="man/figures/screencast.png" width="400"></a>

## Contributing

See our [CONTRIBUTING](https://github.com/ropensci/taxize/blob/master/.github/CONTRIBUTING.md) document.

## Contributors

Collected via GitHub Issues: honors all contributors in alphabetical order. Code contributors are in bold.

[afkoeppel](https://github.com/afkoeppel) - [afredstonhermann](https://github.com/afredstonhermann) - [ahhurlbert](https://github.com/ahhurlbert) - [albnd](https://github.com/albnd) - [Alectoria](https://github.com/Alectoria) - [andzandz11](https://github.com/andzandz11) - **[anirvan](https://github.com/anirvan)** - [antagomir](https://github.com/antagomir) - **[arendsee](https://github.com/arendsee)** - [ArielGreiner](https://github.com/ArielGreiner) - [arw36](https://github.com/arw36) - [ashenkin](https://github.com/ashenkin) - **[ashiklom](https://github.com/ashiklom)** - [benjaminschwetz](https://github.com/benjaminschwetz) - **[benmarwick](https://github.com/benmarwick)** - [bienflorencia](https://github.com/bienflorencia) - [binkySallly](https://github.com/binkySallly) - [bomeara](https://github.com/bomeara) - [BridgettCollis](https://github.com/BridgettCollis) - [bw4sz](https://github.com/bw4sz) - **[cboettig](https://github.com/cboettig)** - [cdeterman](https://github.com/cdeterman) - [ChrKoenig](https://github.com/ChrKoenig) - [chuckrp](https://github.com/chuckrp) - [clarson2191](https://github.com/clarson2191) - [claudenozeres](https://github.com/claudenozeres) - [cmzambranat](https://github.com/cmzambranat) - [cparsania](https://github.com/cparsania) - [daattali](https://github.com/daattali) - [DanielGMead](https://github.com/DanielGMead) - [DarrenObbard](https://github.com/DarrenObbard) - [davharris](https://github.com/davharris) - [davidvilanova](https://github.com/davidvilanova) - [diogoprov](https://github.com/diogoprov) - **[dlebauer](https://github.com/dlebauer)** - [dlenz1](https://github.com/dlenz1) - [dougwyu](https://github.com/dougwyu) - [dschlaep](https://github.com/dschlaep) - **[EDiLD](https://github.com/EDiLD)** - [edwbaker](https://github.com/edwbaker) - [emhart](https://github.com/emhart) - [eregenyi](https://github.com/eregenyi) - [fdschneider](https://github.com/fdschneider) - [fgabriel1891](https://github.com/fgabriel1891) - [fischhoff](https://github.com/fischhoff) - **[fmichonneau](https://github.com/fmichonneau)** - **[fozy81](https://github.com/fozy81)** - **[gedankenstuecke](https://github.com/gedankenstuecke)** - [gimoya](https://github.com/gimoya) - [GISKid](https://github.com/GISKid) - [git-og](https://github.com/git-og) - [glaroc](https://github.com/glaroc) - **[gpli](https://github.com/gpli)** - [gustavobio](https://github.com/gustavobio) - [hlapp](https://github.com/hlapp) - **[ibartomeus](https://github.com/ibartomeus)** - **[Ironholds](https://github.com/Ironholds)** - [jabard89](https://github.com/jabard89) - [jangorecki](https://github.com/jangorecki) - **[jarioksa](https://github.com/jarioksa)** - [jebyrnes](https://github.com/jebyrnes) - **[jeroen](https://github.com/jeroen)** - **[jimmyodonnell](https://github.com/jimmyodonnell)** - [joelnitta](https://github.com/joelnitta) - [johnbaums](https://github.com/johnbaums) - [jonmcalder](https://github.com/jonmcalder) - [jordancasey](https://github.com/jordancasey) - **[josephwb](https://github.com/josephwb)** - [jsgosnell](https://github.com/jsgosnell) - [JulietteLgls](https://github.com/JulietteLgls) - **[jwilk](https://github.com/jwilk)** - [kamapu](https://github.com/kamapu) - **[karthik](https://github.com/karthik)** - **[katrinleinweber](https://github.com/katrinleinweber)** - [KevCaz](https://github.com/KevCaz) - [kgturner](https://github.com/kgturner) - [kmeverson](https://github.com/kmeverson) - [Koalha](https://github.com/Koalha) - **[ljvillanueva](https://github.com/ljvillanueva)** - **[maelle](https://github.com/maelle)** - [Markus2015](https://github.com/Markus2015) - [matutosi](https://github.com/matutosi) - [mcsiple](https://github.com/mcsiple) - [MikkoVihtakari](https://github.com/MikkoVihtakari) - [millerjef](https://github.com/millerjef) - [miriamgrace](https://github.com/miriamgrace) - [MK212](https://github.com/MK212) - [mpnelsen](https://github.com/mpnelsen) - [MUSEZOOLVERT](https://github.com/MUSEZOOLVERT) - [nate-d-olson](https://github.com/nate-d-olson) - [nmatzke](https://github.com/nmatzke) - [npch](https://github.com/npch) - [ocstringham](https://github.com/ocstringham) - [p-neves](https://github.com/p-neves) - [p-schaefer](https://github.com/p-schaefer) - [padpadpadpad](https://github.com/padpadpadpad) - [paternogbc](https://github.com/paternogbc) - **[patperu](https://github.com/patperu)** - [pederengelstad](https://github.com/pederengelstad) - [philippi](https://github.com/philippi) - [Phylloxera](https://github.com/Phylloxera) - **[pmarchand1](https://github.com/pmarchand1)** - [pozsgaig](https://github.com/pozsgaig) - [PrincessPi314](https://github.com/PrincessPi314) - [pssguy](https://github.com/pssguy) - **[raredd](https://github.com/raredd)** - [rec3141](https://github.com/rec3141) - **[Rekyt](https://github.com/Rekyt)** - [RodgerG](https://github.com/RodgerG) - [rossmounce](https://github.com/rossmounce) - [sariya](https://github.com/sariya) - [sastoudt](https://github.com/sastoudt) - [scelmendorf](https://github.com/scelmendorf) - **[sckott](https://github.com/sckott)** - [SimonGoring](https://github.com/SimonGoring) - [snsheth](https://github.com/snsheth) - [snubian](https://github.com/snubian) - [Squiercg](https://github.com/Squiercg) - [sunray1](https://github.com/sunray1) - **[taddallas](https://github.com/taddallas)** - [tdjames1](https://github.com/tdjames1) - [tmkurobe](https://github.com/tmkurobe) - [toczydlowski](https://github.com/toczydlowski) - [tpaulson1](https://github.com/tpaulson1) - [tpoisot](https://github.com/tpoisot) - **[TrashBirdEcology](https://github.com/TrashBirdEcology)** - **[trvinh](https://github.com/trvinh)** - **[vijaybarve](https://github.com/vijaybarve)** - [wcornwell](https://github.com/wcornwell) - [willpearse](https://github.com/willpearse) - [wpetry](https://github.com/wpetry) - [yhg926](https://github.com/yhg926) - **[zachary-foster](https://github.com/zachary-foster)**

## Road map

Check out our [milestones](https://github.com/ropensci/taxize/milestones) to see what we plan to get done for each version.

## Meta

* Please [report any issues or bugs](https://github.com/ropensci/taxize/issues).
* License: MIT
* Get citation information for `taxize` in R doing `citation(package = 'taxize')`
* Please note that this package is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

[![rofooter](https://ropensci.org/public_images/github_footer.png)](https://ropensci.org)