### 2020-10-29 README.Rmd
REVDEPCHECK - remove contribs from readme for cran push

taxize
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![cran checks](https://cranchecks.info/badges/worst/taxize)](https://cranchecks.info/pkgs/taxize)
[![R-CMD-check](https://github.com/ropensci/taxize/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/taxize/actions/)
[![codecov](https://codecov.io/gh/ropensci/taxize/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/taxize)
[![rstudio mirror downloads](https://cranlogs.r-pkg.org/badges/taxize)](https://github.com/r-hub/cranlogs.app)
[![cran version](https://www.r-pkg.org/badges/version/taxize)](https://cran.r-project.org/package=taxize)

`taxize` allows users to search over many taxonomic data sources for species names (scientific and common) and download up and downstream taxonomic hierarchical information - among other things.

The taxize book: https://taxize.dev

Package documentation: https://docs.ropensci.org/taxize/

## Installation

### Stable version from CRAN

```{r eval=FALSE}
install.packages("taxize")
```

### Development version from GitHub

Windows users install Rtools first.

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("ropensci/taxize")
```

```{r}
library('taxize')
```

## Screencast

<a href="https://vimeo.com/92883063"><img src="man/figures/screencast.png" width="400"></a>

## Contributing

See our [CONTRIBUTING](https://github.com/ropensci/taxize/blob/master/.github/CONTRIBUTING.md) document.

## Road map

Check out our [milestones](https://github.com/ropensci/taxize/milestones) to see what we plan to get done for each version.

## Meta

* Please [report any issues or bugs](https://github.com/ropensci/taxize/issues).
* License: MIT
* Get citation information for `taxize` in R doing `citation(package = 'taxize')`
* Please note that this package is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

[![rofooter](https://ropensci.org/public_images/github_footer.png)](https://ropensci.org)