### 2025-02-06 README.Rmd
remove more urls for CRAN checks

---
output: github_document
---

taxize
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![rworkflows](https://github.com/ropensci/taxize/workflows/rworkflows/badge.svg)](https://github.com/ropensci/taxize/actions/)
[![codecov](https://codecov.io/gh/ropensci/taxize/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/taxize)
[![rstudio mirror downloads](https://cranlogs.r-pkg.org/badges/taxize)](https://github.com/r-hub/cranlogs.app)
[![cran version](https://www.r-pkg.org/badges/version/taxize)](https://cran.r-project.org/package=taxize)

`taxize` allows users to search over many taxonomic data sources for species names (scientific and common) and download up and downstream taxonomic hierarchical information - among other things.

Package documentation: https://docs.ropensci.org/taxize/

## Installation

### Stable version from CRAN

```{r eval=FALSE}
install.packages("taxize")
```

### Development version from GitHub

Windows users install Rtools first.

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("ropensci/taxize")
```

```{r}
library('taxize')
```

## Screencast

<a href="https://vimeo.com/92883063"><img src="man/figures/screencast.png" width="400"></a>

## Contributing

See our [CONTRIBUTING](https://github.com/ropensci/taxize/blob/master/.github/CONTRIBUTING.md) document.

## Contributors

Collected via GitHub Issues: honors all contributors in alphabetical order. Code contributors are in bold.

afkoeppel - ahhurlbert - albnd - Alectoria - andzandz11 - **anirvan** - antagomir - **arendsee** - ArielGreiner - arw36 - ashenkin - **ashiklom** - benjaminschwetz - **benmarwick** - bienflorencia - binkySallly - bomeara - BridgettCollis - bw4sz - **cboettig** - cdeterman - ChrKoenig - chuckrp - clarson2191 - claudenozeres - cmzambranat - cparsania - daattali - DanielGMead - DarrenObbard - davharris - davidvilanova - diogoprov - **dlebauer** - dlenz1 - dougwyu - dschlaep - **EDiLD** - edwbaker - emhart - eregenyi - fdschneider - fgabriel1891 - fischhoff - **fmichonneau** - **fozy81** - **gedankenstuecke** - gimoya - git-og - glaroc - **gpli** - gustavobio - hlapp - **ibartomeus** - **Ironholds** - jabard89 - jangorecki - **jarioksa** - jebyrnes - **jeroen** - **jimmyodonnell** - joelnitta - johnbaums - jonmcalder - jordancasey - **josephwb** - jsgosnell - JulietteLgls - **jwilk** - kamapu - **karthik** - **katrinleinweber** - KevCaz - kgturner - kmeverson - Koalha - **ljvillanueva** - **maelle** - Markus2015 - matutosi - mcsiple - MikkoVihtakari - millerjef - miriamgrace - mpnelsen - MUSEZOOLVERT - nate-d-olson - nmatzke - npch - ocstringham - p-neves - p-schaefer - padpadpadpad - paternogbc - **patperu** - pederengelstad - philippi - Phylloxera - **pmarchand1** - pozsgaig - pssguy - **raredd** - rec3141 - **Rekyt** - RodgerG - rossmounce - sariya - sastoudt - scelmendorf - **sckott** - SimonGoring - snsheth - snubian - Squiercg - sunray1 - **taddallas** - tdjames1 - tmkurobe - toczydlowski - tpaulson1 - tpoisot - **TrashBirdEcology** - **trvinh** - **vijaybarve** - wcornwell - willpearse - wpetry - yhg926 - **zachary-foster**

## Road map

Check out our [milestones](https://github.com/ropensci/taxize/milestones) to see what we plan to get done for each version.

## Meta

* Please [report any issues or bugs](https://github.com/ropensci/taxize/issues).
* License: MIT
* Get citation information for `taxize` in R doing `citation(package = 'taxize')`
* Please note that this package is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

[![rofooter](https://ropensci.org/public_images/github_footer.png)](https://ropensci.org)